### 2015-06-17 table_License_Devices_price.tex
Add minitable with License, Devices and price

\documentclass{article}
\usepackage{array,booktabs,graphicx}% http://ctan.org/pkg/{array,booktabs}

\begin{document}

\begin{table}[ht]

  \centering
  \caption{License, Devices and plataform and prices}
  \renewcommand{\arraystretch}{1.5}% Spread rows out...
  \resizebox{8cm}{!}{
  \begin{tabular}{>{\centering\bfseries}m{1in} >{\centering}m{1in} >{\centering}m{1in} >{\centering\arraybackslash}m{1in}}
    \toprule
      & \textbf{License} & \textbf{Devices} & \textbf{Price} \\
    \midrule
    MUSES & OpenSource &  & Free*  \\
    IBM Mobile Security & Propietary &  &  \\
    Sophos Mobile Control & Propietary & IOS, Android, Windows Phone  &  By Request\\
    Samsung Knox and Blackbery Balance & Propietary & Android Samsung Devices and IOS & 12\$ per year/device \\
    Blackphone & Propietary & Own device & 120\$ per year \\
    Android for Work & Propietary & Android and devices with Chrome Browser &  12\$ per month/user\\
    \bottomrule
  \end{tabular}
  }
\end{table}

\end{document}