### 2015-07-08 table_taxonomy.tex
Ladies and gentlemen, the best tables in the nation for personal enjoyment :neckbeard:

\documentclass{article}
\usepackage{array,booktabs,graphicx,colortbl, pifont}% http://ctan.org/pkg/{array,booktabs}

\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\begin{document}

\begin{table}[ht]

  \centering
%  \renewcommand{\arraystretch}{1.2}% Spread rows out...
  \resizebox{14cm}{!}{
  \begin{tabular}{>{\centering\bfseries}m{1in} >{\centering}m{1in} >{\centering}m{1in} >{\centering}m{1in} >{\centering}m{1in} >{\centering\arraybackslash}m{1in}}
    \toprule
      & \textbf{Online threat detection method} & \textbf{Updatable database information} & \textbf{Standalone} & \textbf{Non-invasive data access} & \textbf{Superuser system control} \\
    \midrule
    MUSES & \cmark & \xmark & & &  \\
     \arrayrulecolor[gray]{0.8}\hline
    IBM Mobile Security & & & & &  \\
     \arrayrulecolor[gray]{0.8}\hline
    Sophos Mobile Control & & & & &\\
     \arrayrulecolor[gray]{0.8}\hline
    Samsung Knox and Blackbery Balance & & & & &\\
     \arrayrulecolor[gray]{0.8}\hline
    Blackphone & & & & &\\
     \arrayrulecolor[gray]{0.8}\hline
    Android for Work & & & & &\\
    \bottomrule
  \end{tabular}
  }
  \caption{Features of the analyzed products considering licenses, type of devices and price.}
\end{table}

\end{document}