### 2015-07-08 table_taxonomy.tex

\documentclass{article}
\usepackage{array,booktabs,graphicx,colortbl, pifont}% http://ctan.org/pkg/{array,booktabs}

\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\nmark}{\scalebox{1.25}{\textbf{-}} }

\begin{document}

\begin{table}[ht]

  \centering
  \renewcommand{\arraystretch}{1.8}% Spread rows out...
%  \resizebox{14cm}{!}{
  \begin{tabular}{>{\centering\bfseries}m{0.8in} >{\centering}m{0.8in} >{\centering}m{0.8in} >{\centering}m{0.8in} >{\centering}m{0.8in} >{\centering\arraybackslash}m{0.8in}}
    \toprule
      & \textbf{Offline threat detection method} & \textbf{Updatable database} & \textbf{Standalone} & \textbf{Non-invasive data access} & \textbf{Superuser system control} \\
    \midrule
    MUSES & \cmark & \cmark & \cmark & \cmark & \cmark  \\
     \arrayrulecolor[gray]{0.8}\hline
    IBM Mobile Security & \xmark & \xmark & \cmark & \xmark & \xmark  \\
     \arrayrulecolor[gray]{0.8}\hline
    Sophos Mobile Control & \xmark & \xmark & \cmark & \nmark & \nmark \\
     \arrayrulecolor[gray]{0.8}\hline
    Samsung Knox and Blackbery Balance & \nmark & \nmark & \cmark & \nmark & \nmark\\
     \arrayrulecolor[gray]{0.8}\hline
    Blackphone & \nmark & \nmark & \cmark & \nmark & \nmark\\
     \arrayrulecolor[gray]{0.8}\hline
    Android for Work & \nmark & \nmark & \cmark & \nmark & \nmark\\
    \bottomrule
  \end{tabular}
%  }
  \caption{Proposed taxonomy to classify the BYOD tools described in this paper.}
\end{table}

\end{document}