### 2015-07-08 table_License_Devices_price.tex
Completing @deantares table

\documentclass{article}
\usepackage{array,booktabs,graphicx,colortbl,pifont}% http://ctan.org/pkg/{array,booktabs}

\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\begin{document}

\begin{table}[ht]

  \centering
  \renewcommand{\arraystretch}{1.8}% Spread rows out...
%  \resizebox{8cm}{!}{
  \begin{tabular}{>{\centering\bfseries}m{1in} >{\centering}m{1in} >{\centering}m{1in} >{\centering\arraybackslash}m{1in}}
       \toprule
      & \textbf{License} & \textbf{Devices} & \textbf{Price} \\
       \midrule
    MUSES & OpenSource &  & Free*  \\
       \arrayrulecolor[gray]{0.8}\hline
    IBM Mobile Security & Proprietary & IOS, Android, Windows Phone, Blackberry  &  By Request\\
        \arrayrulecolor[gray]{0.8}\hline
    Sophos Mobile Control & Proprietary & IOS, Android, Windows Phone  &  By Request\\
        \arrayrulecolor[gray]{0.8}\hline
    Samsung Knox and Blackbery Balance & Proprietary & Android Samsung Devices and IOS & 12\$ per year/device \\
        \arrayrulecolor[gray]{0.8}\hline
    Blackphone & Proprietary & Own device & 120\$ per year \\
        \arrayrulecolor[gray]{0.8}\hline
    Android for Work & Proprietary & Android and devices with Chrome Browser &  12\$ per month/user\\
       \bottomrule
  \end{tabular}
%  }
  \caption{Features of the analyzed products considering licenses, type of devices and price.}
\end{table}

\end{document}